---
title: "Re-Analysis of Najarro et al. 2015"
author: "Michael Najarro"
date: "6/21/2020"
output: pdf_document
---

#*Introduction*

In 2015, Najarro et al. produced a publication that partially validated the results of Sharon et al. 2010, which demonstrated a non-genetic form of inheritance for sexual preferences mediated throghu diet and gut-microbiota. Najarro partially replciated the results of Sharon et al. but only in a line of inbred fruit flies.

# *Objective*

This report attempts to demonstrate the findings of Najarro in a reproducible and easily interpretable format. 
The resutls represented are a part of M. Najarro's Masters thesis **The effect of diet treatment on mating preference in Drosophila melanogaster** (Najarro 2015), and the subsequent publication that developed from the thesis, **Choosing mates based on the diet of your ancestors: replication of non-genetic assortative mating in Drosophila melanogaster** (Najarro et al. 2015).

For both document, there were four main objectives:

    1. Validate the assortative mating results of Sharon et al. (2010) by following their experimental protocol.
    
    2. Determine if the phenomenon observed by Sharon et al. (2010) was limited to the genotype of the inbred D. melanogaster strain they used.
    
    3. Build upon their results by performing alternative mate choice design experiments called three-way mate choice experiments, to test the generality of the assortative mating patterns
    
    4. Determine if one sex drove mate preference.



## Step 1: Load data
```{r}
library(pacman)
p_load(readxl, tidyverse, magrittr)
```



```{r}
fourway <- read.csv("./data/Data_from_4-way_mate_choice.csv", stringsAsFactors = FALSE)

threeway <- read.csv("./data/Data_from_3-way_mate_choice.csv", stringsAsFactors = FALSE)
```


These statistics can be found in Najarro 2015  . Each section will identify the source location of the code.



## Step 2: assortative info for canton-S columntable 1
```{r}

###all associative matings
assortative_matings<-nrow(subset(fourway,fourway[,15]== "assortative"))

####assortative matings for canton-s
allcantonsassortativematimgs<-nrow(filter(fourway, fourway[,15]== "assortative", fourway[,2]=="Canton-S"))

####cantonscmy assortative matings
allcmycantonsassortative<-nrow(filter(fourway,fourway[,16]=="CMY", fourway[,17]=="CMY", fourway[,2]=="Canton-S"))
####cantons starch
allcantonstarchassortativecantons<-nrow(filter(fourway,fourway[,16]=="Starch", fourway[,17]=="Starch", fourway[,2]=="Canton-S"))

#filter(fourway,fourway[,16]=="CMY", fourway[,17]=="CMY", fourway[,2]=="allRAL")
#filter(fourway,fourway[,16]=="Starch", fourway[,17]=="Starch", fourway[,2]=="allRAl")
#filter(fourway, fourway[,15]== "assortative", fourway[,2]=="allRAL")
numberofcantos<-nrow(filter(fourway, fourway[,2]=="Canton-S"))

```

###step2disassortative cantons code table 1
```{r}
#disassortative_matings<-nrow(subset(fourway,fourway[,15]== "disassortative"))
allcantonsdisassortativematimgs<-nrow(filter(fourway, fourway[,15]== "disassortative", fourway[,2]=="Canton-S"))
####cantonscmy disassortative matings
cmycantonsdisassortative<-nrow(filter(fourway,fourway[,16]=="CMY", fourway[,17]=="Starch", fourway[,2]=="Canton-S"))
####cantons starch disassortative
cantonstarchdisassortative<-nrow(filter(fourway,fourway[,16]=="Starch", fourway[,17]=="CMY", fourway[,2]=="Canton-S"))

```

### step 3:
```{r}
#binom test for assort matings for cantons 
propassortcantons<-allcantonsassortativematimgs /(allcantonsassortativematimgs+ allcantonsdisassortativematimgs)
#binom.test(x=allcantonsassortativematimgs, n=344,p=0.5 )
cantos4waybinom<-binom.test(x=allcantonsassortativematimgs, n=numberofcantos,p=0.5 )
```
##step5 : assortative info for all ral column table 1
```{r}
numberofallrall<-nrow(filter(fourway, fourway[,2]=="allRAL"))

####assortative matings for all ral
allralassortativematimgs<-nrow(filter(fourway, fourway[,15]== "assortative", fourway[,2]=="allRAL"))

####cantonscmy assortative matings
allralcmyassortative<-nrow(filter(fourway,fourway[,16]=="CMY", fourway[,17]=="CMY", fourway[,2]=="allRAL"))
####cantons starch
allralstarchassortativecantons<-nrow(filter(fourway,fourway[,16]=="Starch", fourway[,17]=="Starch", fourway[,2]=="allRAL"))
```

##step6 : disassort info for all ral column table 1
```{r}
#disassortative_matings<-nrow(subset(fourway,fourway[,15]== "disassortative"))
allralldisassortativematimgs<-nrow(filter(fourway,  fourway[,2]=="allRAL", fourway[,15]== "disassortative"))
####cantonscmy disassortative matings
allrallcmydisassortative<-nrow(filter(fourway, fourway[,2]=="allRAL",fourway[,16]=="CMY", fourway[,17]== "Starch" ))
####cantons starch disassortative
allralstarchdisassortative<-nrow(filter(fourway,fourway[,16]=="Starch", fourway[,17]=="CMY", fourway[,2]=="allRAL"))
```
 ## step 7 Binomial for All Rall:
```{r}

proportionassorttallrall<-allralassortativematimgs/(allralldisassortativematimgs+ allralassortativematimgs)
h<-allralassortativematimgs+allralldisassortativematimgs
#binom.test(x=allralassortativematimgs, n=302,p=0.5 )
allrall4waybinom<-binom.test(x=allralassortativematimgs, n=h,p=0.5 )
allrall4waybinom[3]
```
##Step 8. Oregon-R
```{r}
oregonrmatings<-385
oregonrassortmatings<-236
oregonrdisassortative<-149
oregonrstarchxstarch<-109
oregonrcmyxcmy<-127
oregonrcmyxstarch<-81
oregonrstarchxcmy<-68
propassortoregonr<-oregonrassortmatings/oregonrmatings
oregon4way<-binom.test(x=oregonrassortmatings , n=oregonrmatings, p=0.5)
as.numeric(oregon4way[3])
```

##step9.table1
```{r}
table1<-tribble(~CantonS, ~AllRal, ~OregonR ,
numberofcantos, numberofallrall, oregonrmatings,
allcantonsassortativematimgs,allralassortativematimgs,oregonrassortmatings,
allcantonsdisassortativematimgs,allralldisassortativematimgs,oregonrdisassortative,
propassortcantons,proportionassorttallrall,propassortoregonr,
as.numeric(cantos4waybinom[3]),as.numeric(allrall4waybinom[3]),as.numeric(oregon4way[3]))
tableone<-as.matrix(table1)
rownames(tableone)<-c( "N matings", "assortativ_matings","disassortative_matings","proportion assortative matings","pvalues of assortative matings in binomial test.")
colnames(tableone)<- c("Cantons", "AllRal","OregonR")

tableone

```
##step 10. new table 1.5 of assortative matings  matings
```{r}
assortative<-tribble(~CantonS, ~AllRal, ~OregonR,
        allcantonstarchassortativecantons,allralstarchassortativecantons,oregonrstarchxstarch, allcmycantonsassortative,allralcmyassortative,oregonrcmyxcmy)
as<- as.data.frame(assortative)  
as
row.names(as)<-c("starchassortative matings","cmy assortative")      
as
```

##step10 new table1.75 of disassortative   matings
```{r}
dis<-tribble(~CantonS, ~AllRal, ~OregonR,
        cantonstarchdisassortative ,allralstarchdisassortative,oregonrdisassortative,
        cmycantonsdisassortative, allrallcmydisassortative,oregonrcmyxstarch,
        )
disa<-as.data.frame(dis)
rownames(disa)<-c( "disassortativematings with starch rearted females", "disassortative matings with a cmy femalereared")
colnames(table)<- c("Cantons", "AllRal","OregonR")
       
```
##table2 stats three way statistics
```{r}

###asassortative matings with male cmy chooser
MC <-nrow(filter(threeway,threeway[,15]== "assortative", threeway[,2 ]=="Canton s",threeway[,16]=="cmy", threeway[,6]=="MC" ))
 
#####female cmy chooser
 FC<-nrow(filter(threeway,threeway[,15]== "assortative", threeway[,2 ]=="Canton s", threeway[,6]=="FC" ))

 ####assortative starch matings where the chooser was male and raised on starch
 ms<-nrow(filter(threeway,threeway[,15]== "assortative", threeway[,2 ]=="Canton s", threeway[,6]=="MS" ))
 
 ####assortative starch matings where the chooser was female raised on starch
 fs<-nrow(filter(threeway,threeway[,15]== "assortative", threeway[,2 ]=="Canton s", threeway[,6]=="FS" ))

####all assortative matings for three way matings
#nrow(filter(threeway,threeway[,15]== "assortative", threeway[,2 ]=="Canton s",threeway[,16]=="starch" ))

 

####chooser is malestarch and outcome is disassortative
dismsrow<-nrow(filter(threeway,threeway[,15]== "disassortative", threeway[,2 ]=="Canton s",threeway[,6]=="MS"))
####chooser is female starch and outcome is disassortative
disfsrow<-nrow(filter(threeway,threeway[,15]== "disassortative", threeway[,2 ]=="Canton s",threeway[,6]=="FS"  ))

####chooser is malestarch and outcome is disassortative
dismcrow <-nrow(filter(threeway,threeway[,15]== "disassortative", threeway[,2 ]=="Canton s",threeway[,6]=="MC"))
####chooser is female starch and outcome is disassortative
disfc<-nrow(filter(threeway,threeway[,15]== "disassortative", threeway[,2 ]=="Canton s",threeway[,6]=="FC"  ))
```
#build table 2.
```{r}
column1<- c(fs,ms,FC,MC)
column2<- c(disfsrow,dismsrow, disfc, dismcrow)
column3<- column1+ column2
#column4<- c(column1/column3)
a<-column1[1]
v<-column3[1]
test1<-binom.test(x=a, n=v,p=0.5)
q<- column1[2]
w<- column3[2]
test2<-binom.test(x=q, n=w,p=0.5)
r<- column1[3]
t<-column3[3]
test3<- binom.test(x=r, n=t,p=0.5)
y<-column1[4]
u<-column3[4]
test4<- binom.test(x=y,n=u,p=0.5)
column4<- c(test1[3], test2[3],test4[3])
table2<-data.frame(column1, column2, column3,column4)
colnames(table2)<-c("assortative matings","disassortative matings","sum of assortative and disassortative matings for a row"," binomial test for assortative matings,p value ")
rownames(table2)<- c("female starch chooser",
                     "male starch chooser",
                     "female cmy chooser",
                     "male cmy chooser")
```

```{r}
#figure1
proportionassortativematingsbygenotype<- c( proportionassorttallrall, propassortcantons, propassortoregonr)
strains<- c("assort all rall","assortative Cantons","Oregon- R")
pr<-data.frame(proportionassortativematingsbygenotype, strains)
pr
ggplot(pr, aes(x=strains, y= proportionassortativematingsbygenotype) )+geom_point()
```
```{r}
##anova models to evalate influence of other variables

```


